\documentclass[12pt]{extarticle}
\usepackage[utf8]{inputenc}
\usepackage{cite}

\title{Cover letter}
\date{}

\begin{document}

\maketitle

Dear Dr. Len,\\

My name is Thibault Poiret and I would like to apply for the postdoctoral position you advertised, related to constructing moduli spaces for Prym varieties and their tropicalizations.

I completed my PhD in algebraic geometry at the universities of Bordeaux and Leiden under the supervision of Prof. Qing Liu and Prof. Bas Edixhoven, working on Néron models over bases of arbitrary dimension, with a particular emphasis on families of prestable curves. I am now a postdoctoral fellow in Cambridge, in the group of Dr. Dhruv Ranganathan. Prestable curves, their tropicalizations and their moduli have been the most central object in my research so far; and I think my experience in working on the logarithmic and tropical Jacobians will be particularly relevant in thinking about the moduli of Prym varieties. This is a topic I have already had some interesting conversations about, and I would like to explore it more seriously.

In the past, I have worked both on autonomous and collaborative projects. Exchanging ideas with others is my favourite part of the job. As a consequence, I am enthusiastic about being an active participant in seminars and working groups. In research as in teaching, I treat the question of how to communicate ideas in the clearest way possible with the seriousness it deserves.


I hope you will consider this application, and give me this opportunity to learn many new and interesting things.\\

Best regards,

Thibault Poiret

\end{document}